---
title: "team_04_tidyverse.qmd"
format: 
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    warning: false
author: "Ashsyahid, Owen, Sherwyn, Xuan Yu, Nigel"
number_sections: false
knitr:
  opts_chunk:
    message: false
---

```{r}
#| label: load-packages
#| include: false

library(gt)
library(tidyverse)
```

# Task 1: Converting EPL Football Results into a League Table

## 1.1 Generate a League Table from the EPL Results

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(gt)
```

```{r warning=FALSE, message=FALSE}
epl_results <- read_csv("epl/epl_results.csv")
epl_teams <- read_csv("epl/epl_teams.csv")
```

```{r}
# Prepare the results data
epl_results_long <- epl_results |>
  pivot_longer(cols = -home_team, names_to = "away_team", values_to = "result") |>
  separate(result, into = c("home_goals", "away_goals"), sep = "-", convert = TRUE) |>
  filter(!is.na(home_goals))
```

```{r}
# Calculate home team statistics
home_stats <- epl_results_long |>
  group_by(home_team) |>
  summarize(
    points_home = sum(if_else(home_goals > away_goals, 3, if_else(home_goals == away_goals, 1, 0))),
    goals_for_home = sum(home_goals),
    goals_against_home = sum(away_goals)
  )
```

```{r}
# Calculate away team statistics
away_stats <- epl_results_long |>
  group_by(away_team) |>
  summarize(
    points_away = sum(if_else(away_goals > home_goals, 3, if_else(away_goals == home_goals, 1, 0))),
    goals_for_away = sum(away_goals),
    goals_against_away = sum(home_goals)
  )
```

```{r}
# Combine home and away statistics
team_stats <- home_stats |>
  rename(team = home_team) |>
  full_join(away_stats, by = c("team" = "away_team")) |>
  mutate(
    points = points_home + points_away,
    goals_for = goals_for_home + goals_for_away,
    goals_against = goals_against_home + goals_against_away,
    goal_diff = goals_for - goals_against
  ) |>
  select(team, points, goal_diff, goals_for)
```

```{r}
# Add full team names
league_table <- team_stats |>
  left_join(epl_teams, by = c("team" = "abbr")) |>
  select(team = name, points, goal_diff, goals_for)
```

```{r}
# Rank the teams
league_table <- league_table |>
  arrange(desc(points), desc(goal_diff), desc(goals_for)) |>
  mutate(rank = row_number()) |>
  select(rank, team, points, goal_diff, goals_for)
```

```{r}
# Display the league table
league_table
```
...

## 1.2 Improve Visualization Using `gt` Functions and Discuss rationale
```{r}
# Create a gt table
gt_table <- league_table |>
  gt() |>
  tab_header(
    title = md("**English Premier League Table**"),
    subtitle = md("Season: 2023/2024")
  ) |>
  cols_label(
    goal_diff = "goal difference",
    goals_for = "goals scored"
  ) |>
  fmt_number(
    columns = vars(points, goal_diff, goals_for),
    decimals = 0
  ) |>
  data_color(
    columns = vars(points),
    colors = scales::col_numeric(
      palette = c("red", "white", "green"),
      domain = c(min(league_table$points), max(league_table$points))
    )
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) |>
  tab_style(
    style = cell_borders(sides = "all", color = "grey90", weight = px(1)),
    locations = cells_body(everything())
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "#f7f7f7"),
      cell_text(color = "black")
    ),
    locations = cells_body(columns = vars(rank, team))
  ) |> 
  tab_options(
    table.font.names = "Arial",
    table.font.size = px(12),
    heading.title.font.size = px(16),
    heading.subtitle.font.size = px(14)
  ) |>
  tab_source_note("Source: English Premier League") |>
  tab_style(
    cell_fill(color = "steelblue1"),
    locations = cells_body(
      columns = rank,
      rows = 1:4
    )
  ) |>
  tab_style(
    cell_fill(color = "red"),
    locations = cells_body(
      columns = rank,
      rows = 18:20
    )
  ) |>
  tab_options(table.font.size = px(15))

# Display the table
gt_table
```
**Rationale:**
We used the `gt` package to create a more visually appealing and interactive table for the EPL league table. The `gt` package provides a wide range of customization options for tables, allowing us to improve the readability and aesthetics of the league table. We added styling, formatting, and color-coding to enhance the presentation of the data. The table now includes bold headers, color-coded points column, and custom styling for the top and bottom teams to make them stand out. The source note was also added for transparency and credibility. In the "rank" column for the top 4 teams, we used a light blue color to highlight their qualifications for the Champions League. As for the bottom 3 teams, we used red to highlight their relegation status. In the "points" column, we used a color gradient from red to green to represent the points range. These visual cues help users quickly identify key information and make the table more engaging and informative.

# Task 2: Finding and Visualizing a Real-World Dataset

...

# Task 3: Visualizing Michelson’s Speed-of-Light Measurements

```{r}
#| echo: true
#| eval: true
data(morley)
morley <- as_tibble(morley)
morley
```
## 3.1 Convert Expt Column to Factor

```{r}
#| eval: true
#| echo: true

morley <- morley |>
  mutate(Expt = as.factor(Expt))
morley
```
## 3.2 Plot 1 – Create a Scatter Plot displaying Experiment Number vs. Measured Speeds

```{r}
#| eval: true
#| echo: true

morley |>
  ggplot(aes(x = Expt, y = Speed)) +
  geom_point() +
  labs(title = "Experiment Number vs. Measured Speeds",
       caption = "Source: R data set `morley`",
       x = "Experiment Number",
       y = "Measured Speeds")
```
## 3.3 Plot 2 – Create a Box Plot Measured Speeds with One Box Experiment

```{r}
#| eval: true
#| echo: true

morley |>
  ggplot(aes(x = Expt, y = Speed)) +
  geom_boxplot() +
  labs(title = "Measured Speeds with One Box Experiment",
       caption = "Source: R data set `morley`",
       x = "Experiment Number",
       y = "Measured Speeds")
```
## 3.4 Plot 3 – Add a Violin Plot in Plot 2

```{r}
#| eval: true
#| echo: true

morley |>
  ggplot(aes(x = Expt, y = Speed)) +
  geom_violin(fill = "lightblue", alpha = 0.3) +
  geom_boxplot(width = 0.2) +
  labs(title = "Measured Speeds with One Box Experiment with Violin plots",
       caption = "Source: R data set `morley`",
       x = "Experiment Number",
       y = "Measured Speeds")
```

## 3.5 Plot 4 – Create a Faceted Plot Containing 5 Histograms

```{r}
#| eval: true
#| echo: true

morley |>
  ggplot(aes(x = Speed)) +
  geom_histogram(bins = 10, fill = NA, color="black") +
  facet_wrap(~Expt) +
  labs(title = "Faceted Plot Containing 5 Histograms",
       caption = "Source: R data set `morley`",
       x = "Measured Speeds",
       y = "Frequency")
```
## 3.6 Discuss Your Choice Among the 4 Visualization Plots

We choose Plot 3 - Violin Plot + Box Plot. Since violin plot can be used for grouped data, this plot is chosen because it provides a good visualization of the distribution of the measured speeds across the different experiments. The box plot provides a summary of the distribution of the measured speeds, while the violin plot provides a more detailed view of the distribution of the measured speeds. The combination of the two plots allows us to see the spread of the measured speeds and the distribution of the measured speeds within each experiment.   
