---
title: "week_10_lab_personal"
author: "Ashsyahid, Owen, Sherwyn, Xuan Yu, Nigel"
date: "2024-07-09"
date-format: "dddd MMM D, YYYY"
format:
  html:
    toc: true
    toc-depth: 3
    toc_location: left
    number-sections: false
    number-depth: 3
    embed-resources: true
---

# Task 1: Converting EPL Football Results into a League Table

## 1.1 Generate a League Table from the EPL Results

generate league table tibble with the following columns:
rank team                    points goal_diff goals_for

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(gt)
```

```{r warning=FALSE, message=FALSE}
epl_results <- read_csv("epl/epl_results.csv")
epl_teams <- read_csv("epl/epl_teams.csv")
```

```{r}
# Prepare the results data
epl_results_long <- epl_results %>%
  pivot_longer(cols = -home_team, names_to = "away_team", values_to = "result") %>%
  separate(result, into = c("home_goals", "away_goals"), sep = "-", convert = TRUE) %>%
  filter(!is.na(home_goals))
```

```{r}
# Calculate home team statistics
home_stats <- epl_results_long %>%
  group_by(home_team) %>%
  summarize(
    points_home = sum(if_else(home_goals > away_goals, 3, if_else(home_goals == away_goals, 1, 0))),
    goals_for_home = sum(home_goals),
    goals_against_home = sum(away_goals)
  )
```

```{r}
# Calculate away team statistics
away_stats <- epl_results_long %>%
  group_by(away_team) %>%
  summarize(
    points_away = sum(if_else(away_goals > home_goals, 3, if_else(away_goals == home_goals, 1, 0))),
    goals_for_away = sum(away_goals),
    goals_against_away = sum(home_goals)
  )
```

```{r}
# Combine home and away statistics
team_stats <- home_stats %>%
  rename(team = home_team) %>%
  full_join(away_stats, by = c("team" = "away_team")) %>%
  mutate(
    points = points_home + points_away,
    goals_for = goals_for_home + goals_for_away,
    goals_against = goals_against_home + goals_against_away,
    goal_diff = goals_for - goals_against
  ) %>%
  select(team, points, goal_diff, goals_for)
```

```{r}
# Add full team names
league_table <- team_stats %>%
  left_join(epl_teams, by = c("team" = "abbr")) %>%
  select(team = name, points, goal_diff, goals_for)
```

```{r}
# Rank the teams
league_table <- league_table %>%
  arrange(desc(points), desc(goal_diff), desc(goals_for)) %>%
  mutate(rank = row_number()) |>
  select(rank, team, points, goal_diff, goals_for)
```

```{r}
# Display the league table
league_table
```
...

## 1.2 Improve Visualization Using `gt` Functions
```{r}
# Create a gt table
gt_table <- league_table |>
  gt() |>
  tab_header(
    title = md("**English Premier League Table**"),
    subtitle = md("Season: 2023/2024")
  ) |>
  cols_label(
    goal_diff = "goal difference",
    goals_for = "goals scored"
  ) |>
  fmt_number(
    columns = vars(points, goal_diff, goals_for),
    decimals = 0
  ) |>
  data_color(
    columns = vars(points),
    colors = scales::col_numeric(
      palette = c("red", "white", "green"),
      domain = c(min(league_table$points), max(league_table$points))
    )
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) |>
  tab_style(
    style = cell_borders(sides = "all", color = "grey90", weight = px(1)),
    locations = cells_body(everything())
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "#f7f7f7"),
      cell_text(color = "black")
    ),
    locations = cells_body(columns = vars(rank, team))
  ) |> 
  tab_options(
    table.font.names = "Arial",
    table.font.size = px(12),
    heading.title.font.size = px(16),
    heading.subtitle.font.size = px(14)
  ) |>
  tab_source_note("Source: English Premier League") |>
  tab_style(
    cell_fill(color = "steelblue1"),
    locations = cells_body(
      columns = rank,
      rows = 1:4
    )
  ) |>
  tab_style(
    cell_fill(color = "red"),
    locations = cells_body(
      columns = rank,
      rows = 18:20
    )
  ) |>
  tab_options(table.font.size = px(15))

# Display the table
gt_table
```
**Rationale:**
We used the `gt` package to create a more visually appealing and interactive table for the EPL league table. The `gt` package provides a wide range of customization options for tables, allowing us to improve the readability and aesthetics of the league table. We added styling, formatting, and color-coding to enhance the presentation of the data. The table now includes bold headers, color-coded points column, and custom styling for the top and bottom teams to make them stand out. The source note was also added for transparency and credibility. In the "rank" column for the top 4 teams, we used a light blue color to highlight their qualifications for the Champions League. As for the bottom 3 teams, we used red to highlight their relegation status. In the "points" column, we used a color gradient from red to green to represent the points range. These visual cues help users quickly identify key information and make the table more engaging and informative.

# Task 2: Finding and Visualizing a Real-World Dataset

...

# Task 3: Visualizing Michelson’s Speed-of-Light Measurements

...