---
title: "team_04_tidyverse.qmd"
format: 
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
author: "Ashsyahid, Owen, Sherwyn, Xuan Yu, Nigel"
number_sections: false
knitr:
  opts_chunk:
    message: false
---

```{r}
#| label: load-packages
#| include: false

library(gt)
library(tidyverse)
```

# Task 1: Converting EPL Football Results into a League Table

## 1.1 Generate a League Table from the EPL Results

...

## 1.2 Improve Visualization Using `gt` Functions

...

# Task 2: Finding and Visualizing a Real-World Dataset

...

# Task 3: Visualizing Michelson’s Speed-of-Light Measurements

```{r}
#| echo: true
#| eval: true
data(morley)
morley <- as_tibble(morley)
morley
```
## 3.1 Convert Expt Column to Factor

```{r}
#| eval: true
#| echo: true

morley <- morley %>%
  mutate(Expt = as.factor(Expt))
morley
```
## 3.2 Plot 1 – Create a Scatter Plot displaying Experiment Number vs. Measured Speeds

```{r}
#| eval: true
#| echo: true

morley %>%
  ggplot(aes(x = Expt, y = Speed)) +
  geom_point() +
  labs(title = "Experiment Number vs. Measured Speeds",
       caption = "Source: R data set `morley`",
       x = "Experiment Number",
       y = "Measured Speeds")
```
## 3.3 Plot 2 – Create a Box Plot Measured Speeds with One Box Experiment

```{r}
#| eval: true
#| echo: true

morley %>%
  ggplot(aes(x = Expt, y = Speed)) +
  geom_boxplot() +
  labs(title = "Measured Speeds with One Box Experiment",
       caption = "Source: R data set `morley`",
       x = "Experiment Number",
       y = "Measured Speeds")
```
## 3.4 Plot 3 – Add a Violin Plot in Plot 2

```{r}
#| eval: true
#| echo: true

morley %>%
  ggplot(aes(x = Expt, y = Speed)) +
  geom_violin(fill = "lightblue", alpha = 0.3) +
  geom_boxplot(width = 0.2) +
  labs(title = "Measured Speeds with One Box Experiment with Violin plots",
       caption = "Source: R data set `morley`",
       x = "Experiment Number",
       y = "Measured Speeds")
```

## 3.5 Plot 4 – Create a Faceted Plot Containing 5 Histograms

```{r}
#| eval: true
#| echo: true

morley %>%
  ggplot(aes(x = Speed)) +
  geom_histogram(bins = 10, fill = NA, color="black") +
  facet_wrap(~Expt) +
  labs(title = "Faceted Plot Containing 5 Histograms",
       caption = "Source: R data set `morley`",
       x = "Measured Speeds",
       y = "Frequency")
```
## 3.6 Discuss Your Choice Among the 4 Visualization Plots

We choose Plot 3 - Violin Plot + Box Plot. Since violin plot can be used for grouped data, this plot is chosen because it provides a good visualization of the distribution of the measured speeds across the different experiments. The box plot provides a summary of the distribution of the measured speeds, while the violin plot provides a more detailed view of the distribution of the measured speeds. The combination of the two plots allows us to see the spread of the measured speeds and the distribution of the measured speeds within each experiment.   
